Extreme heterogeneity of influenza virus infection in single cells

### 2018-01-11 paper/ResponseToReviews.tex

Then we can somehow work in the Watanabe reference.

We have the oxidative stress thing to add as a figure supplement.
}

4. All analysis of viral gene expression seems to be at the segment level. How do different transcripts expressed from the same gene segment compare (i.e. NS1 and NEP)?

{\color{black}
This is an important point that we failed to adequately discuss in the original manuscript.
As the reviews note, two of the influenza transcripts (M1 / M2  and NS1 / NS2) are generated from alternative splicing of gene segments.
However, each pair of these transcripts share the same 3' end.
Most current single-cell mRNA sequencing strategies, including the 10X platform that we use, can only sequence the 3' end of the transcript.
The reason is that the cell barcode is appended to this end, and so Illumina sequencing can only extend to regions proximal to the 3' end if they are still to capture the barcode (as is required for single-cell mRNA sequencing).
Therefore, these techniques cannot accurately identify alternative splicing -- this is simply an inherent limitation of the current techniques.
We now clearly explain this fact in the following added text:
\begin{quote}
\textsl{
Note that influenza virus expresses ten major gene transcripts from its eight gene segments, as the M and NS segments are alternatively spliced to produce the M1 / M2 and NS1 / NEP transcript, respectively (Dubois et al, 2014).
However, an inherent limitation of current established single-cell mRNA sequencing techniques is that they only sequence the 3' end of the transcript (Zheng et al., 2017; Macosko et al., 2015; Klein et al., 2015; Cao et al, 2017).
Since the alternative spliceoforms M1 / M2 and NS1 / NEP share the same 3' ends, we cannot distinguish them and therefore will refer simply to the combined counts of transcripts from each of these alternatively spliced segments as the M and NS genes.}
\end{quote}
}

@@

-Why was media change skipped in the replicate experiment?

{\color{red} Needs to be explained}

-Consider using different color schemes for Fig 4A and B, as different variables are being shown.

{\color{black} 
This is a good suggestion.
We now use different color schemes for Figures 4A and 4B.
We also changed the color scheme for Figure 4E to be different than those for 4A and 4B since it also shows a different variable.}

### 2018-01-11 paper/paper.tex
explain about inability to distinguish spliced flu genes

We examined the total amount of viral mRNA versus the expression of each viral gene (Figure~\ref{fig:fluburdenbyflugene}A, Figure~\ref{fig:fluburdenbyflugene}-Figure~supplement~\ref{figsupp:fluburdensamplebreakdown}). 
Cells that lack an RNP gene never derive more than a few percent of their mRNAs from virus, confirming the expected result that all four RNP genes are essential for high levels of viral transcription.
However, we observe cells that lack each of the other non-RNP genes but still derive $\approx$40\% of their mRNAs from virus, suggesting that none of the other genes are important for high levels of viral transcription.
These results are statistically supported by Figure~\ref{fig:fluburdenbyflugene}B, which shows that absence of any RNP gene but \emph{not} any other viral gene is associated with reduced amounts of viral mRNA. 
However, gene absence clearly does not explain all of the variability in viral gene expression, since even cells expressing all viral genes exhibit a very wide distribution in the amount of viral mRNA that they express.
Specifically, at both 8 and 10 hours, the amount of viral mRNA in individual cells expressing all eight viral genes still ranges from $<$1\% to $>$50\% (Figure~\ref{fig:fluburdenbyflugene}A, Figure~\ref{fig:fluburdenbyflugene}-Figure~supplement~\ref{figsupp:fluburdensamplebreakdown}).

>>>>

We examined the total amount of viral mRNA versus the expression of the genes from each viral segment (Figure~\ref{fig:fluburdenbyflugene}A, Figure~\ref{fig:fluburdenbyflugene}-Figure~supplement~\ref{figsupp:fluburdensamplebreakdown}). 
Note that influenza virus expresses ten major gene transcripts from its eight gene segments, as the M and NS segments are alternatively spliced to produce the M1 / M2 and NS1 / NEP transcript, respectively~\citep{dubois2014influenza}.
However, an inherent limitation of current established single-cell mRNA sequencing techniques is that they only sequence the 3' end of the transcript~\citep{zheng2017massively,macosko2015highly,Klein:2015ki,cao2017comprehensive}.
Since the alternative spliceoforms M1 / M2 and NS1 / NEP share the same 3' ends, we cannot distinguish them and therefore will refer simply to the combined counts of transcripts from each of these alternatively spliced segments as the M and NS genes.

Cells that lack an RNP gene never derive more than a few percent of their mRNAs from virus, confirming the expected result that all four RNP genes are essential for high levels of viral transcription (Figure~\ref{fig:fluburdenbyflugene}A, Figure~\ref{fig:fluburdenbyflugene}-Figure~supplement~\ref{figsupp:fluburdensamplebreakdown}).
However, we observe cells that lack each of the other non-RNP genes but still derive $\approx$40\% of their mRNAs from virus, suggesting that none of the other genes are important for high levels of viral transcription.
These results are statistically supported by Figure~\ref{fig:fluburdenbyflugene}B, which shows that absence of any RNP gene but \emph{not} any other viral gene is associated with reduced amounts of viral mRNA. 
However, gene absence clearly does not explain all of the variability in viral gene expression, since even cells expressing all viral genes exhibit a very wide distribution in the amount of viral mRNA that they express.
Specifically, at both 8 and 10 hours, the amount of viral mRNA in individual cells expressing all eight viral genes still ranges from $<$1\% to $>$50\% (Figure~\ref{fig:fluburdenbyflugene}A, Figure~\ref{fig:fluburdenbyflugene}-Figure~supplement~\ref{figsupp:fluburdensamplebreakdown}).
